---
title: "Summary of what we learnt among the Rauks<br><br><br><br><br>"
output: 
  ioslides_presentation:
    css: style.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```


## Take home things

Robust and reproducible data analysis through

- Work in projects in RStudio
- Git and GitKraken/Source Tree
- Rmarkdown
- Making functions instead of copy paste


Additional things that help with robustness and reproducibility indirectly

- Naming things clearly and consistently
- Tidyverse

## Benefits from working in projects

- No need for setwd() (or getwd())
- No need to mix up global variables between projects

NOW A DEMO!

## Git and GitHub

- Version control
- Collaborations

NOW A DEMO!

## Rmarkdown

## Making functions

- If you do something more than twice: make it into a function!
- Less error prone
- Cleaner code in actual analysis script
- Keep them short
- Make wrappers for smaller functions for bigger workflow


## Things that are good to remember

Divide and conquer

## Things that opened our minds

- ggplot
- Interactive plots and shiny apps

## Presention

This presention was done using Rmarkdown in RStudio

```{r , echo = TRUE}
sessionInfo()
```

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

