---
title: "Functions for basic QC and visualization of autoimmunity data (rappp)"
author: "Cecilia Hellström"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: TRUE
vignette: >
  %\VignetteIndexEntry{Functions for basic QC and visualization of autoimmunity data (rappp)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=F}
library(rappp)
```

This package includes functions to enable a standardized workflow for quality control and 
some basic analysis and visualizations of autoimmunity data.
It also includes some functions related to peptide design and other handy functions (documentation will be created later).

This vignette will demonstrate a basic workflow using some of the included functions.
Please see the individual function help documentation for more information on arguments etc.

# Input data structure

Most functions are based on the below input structure, which is then built upon by the different functions. 
The list elements are described in more detail below.
The package includes a mock dataset (MockSBA) which is used for the below examples.

```{r, echo=T, eval=F}
        SBA <- list(MFI=assay_mfi,
                    COUNT=bead_count,
                    SAMPLES=sampleinfo,
                    BEADS=beadinfo,
                    CT=test_mfi,
                    FILTERINFO=NULL)
```

```{r, echo=F, eval=T}
SBA <- MockSBA
```

## MFI
The MFI element contains the raw MFI in a data frame, with samples as rows and beads as columns.
It is recommended to use the bead name structure Gene_Identifier as column names. 
For antigens mapping to more than one gene, it is recommended to only display the first gene followed by an asterix.
Make sure the columns match the rows in the BEADs element, and the rows match the rows in the SAMPLES element.

```{r}
SBA$MFI[1:10, 1:8]
```

## COUNT
The COUNT element contains the bead count in a data frame, with samples as rows and beads as columns.
Use same naming and structure as the MFI element.

```{r}
SBA$COUNT[1:10, 1:8]
```

## SAMPLES
The SAMPLES element contains all sample information you wish to use in a data frame, with samples as rows.
Make sure the rows are sorted in the same order as the rows in the MFI and COUNT elements.

```{r}
SBA$SAMPLES[1:10,]
```

## BEADS
The BEADS element contains all bead information you wish to use in a data frame, with beads as rows.
Make sure the rows are sorted in the same order as the columns in the MFI, COUNT and CT elements.

```{r}
SBA$BEADS[1:8, 1:4]
```

## CT
The CT element contains the raw MFI values for the coupling efficiency test in a data frame, 
with replicates meassuring the His6ABP-tag (or other tag) as rows and beads as columns.
Use same column naming and structure as in the MFI and BEADS elements.

```{r}
SBA$CT[, 1:8]
```

# Quality control plot/filtering/flagging functions

## Flagging based on coupling test
The function ap_ct() produces a plot (if you want) in a pdf (if you want), and flags all antigens of the specified types which have a signal lower than the empty bead (or other bead you want to base the flagging threshold on).
The element BEADS in the input list is updated with the flagging annotation and the element FILTERINFO is updated to state that ct-flagging has been performed.

```{r ct, fig.width=13, fig.height=6}
SBA <- ap_ct(SBA, empty_bead=grep("Bare", colnames(SBA$MFI)), 
             empty_co_multiple = 3, types = "PrEST",
  shouldplot = TRUE, shouldpdf = FALSE,useDingbats = FALSE)

# Updated BEADS
SBA$BEADS[which(SBA$BEADS$Flagged != ""), 1:4]

# Updated FILTERINFO
SBA$FILTERINFO
```

## Anti human-IgG, loading control
The function ap_igx() produces a plot (if you want) in a pdf (if you want), and filters all samples with low signal.
You decide if the filtering cutoff should be a fixed value or for example based on the empty wells.
The element SAMPLES in the input list is updated with the filtering annotation and the element FILTERINFO is updated.

```{r ahIgG, fig.width=13, fig.height=8}
SBA <- ap_igx(SBA, IgX_bead=grep("IgG", colnames(SBA$MFI)), IgType = "G", 
              IgX_cutoff = 5000, cosfac = c(3,-3), 
              shouldplot = TRUE, shouldpdf = FALSE, useDingbats = FALSE)

# Updated SAMPLES
SBA$SAMPLES[which(SBA$SAMPLES$Filtered != ""),]

# Updated FILTERINFO
SBA$FILTERINFO
```

## Bead count
The function ap_count() produces a plot (if you want) in a pdf (if you want), and filters or flags samples and beads based on their bead count.
You decide which filtering/flagging cutoffs to use, and how many samples with low bead count that are needed to filter a bead.
The elements SAMPLES and BEADS in the input list are updated with the filtering and falgging annotation and the element FILTERINFO is updated.
In the first plot (samples) nothing is filtered, for the second plot (beads) samples filtered based on bead count are not included, 
and for the last two plots (samples and beads) both samples and beads filtered based on bead count are excluded.

```{r count, fig.width=13, fig.height=8}
SBA <- ap_count(SBA, labels = "Gene_AgID", protein = "GeneShort",
  agID = "PrEST.ID", samp_co = 32, bead_flag = 32, bead_filter = 16,
  N_filter = 0, shouldplot = TRUE, shouldpdf = FALSE, useDingbats = FALSE)

# Updated SAMPLES
SBA$SAMPLES[which(SBA$SAMPLES$Filtered != ""),]

# Updated BEADS
SBA$BEADS[which(SBA$BEADS$Flagged != "" | SBA$BEADS$Filtered != ""), 1:4]

# Updated FILTERINFO
SBA$FILTERINFO
```

## Manual flagging/filtering annotation

You may also want to add some flagging or filtering annotations manually, for example for control beads and wells:
```{r manual}
        SBA$SAMPLES$Filtered[grep("rep|pool|commercial|mix|blank|empty|buffer", SBA$SAMPLES$sample_name, ignore.case=T)] <- "Control"
        SBA$BEADS$Filtered[grep("ebna|ahig|human", SBA$BEADS$Gene_AgID, ignore.case=T)] <- "PosControl"
        SBA$BEADS$Filtered[grep("bare|empty|his6abp|hisabp", SBA$BEADS$Gene_AgID, ignore.case=T)] <- "NegControl"
        SBA$FILTERINFO <- c(SBA$FILTERINFO, "TechControls")
```


# Data transformation (MADs, scores, binary)

The package includes functions that do MAD-transformation, binning (scoring), cutoff selection based on the scoring and create binary tables.
These functions are combined into the wrapper function ap_norm2().
You select which MAD-values you want to use for the binning (scoring).
Since the transformation etc. are done per sample, no samples are excluded.
Antigens without any filtering annotations and with the annotation "NegControl" are included in the transformation etc. 
Several new elements will be added to the input list: MADS, CUTOFF_KEY, SCORE, BINARY, DENS, ANTIGEN_CUTOFFS, and ANTIGEN_CUTOFFS_CONT.

```{r norm}
SBA <- ap_norm2(SBA, MADlimits = seq(0, 70, 5), na.rm = TRUE,
  check.names = FALSE, mad_constant = 1, mad_low = FALSE,
  mad_high = FALSE, score_rightmost.closed = FALSE,
  score_left.open = FALSE, score_all.inside = FALSE,
  coselect_slope_cutoff = -0.5, coselect_offset = 0.1,
  coselect_bw = 0.1)

# Updated SBA
lapply(SBA, class)

```

# Results
The function ap_reactsummary2() calculates number of reactivities and the corresponding reactivity frequencies 
per sample and antigen for each cutoff used in the data transformation. It will also calculate per group, 
either all if no grouping is defined or for a grouping of your choice. 
If more than one group it will also do Fisher's exact test and calculate the frequency difference between each pair of groups. 
This function does not build upon the input list, but has a separate output list, one per grouping.

The function ap_reactsummary2() is incorporated in the function ap_agresults(), which also visualizes the data and reactivities per bead.
The heatmap for Fisher's exact test p-values is only included if there is more than one sample group.

In the below example I've only selected one bead to plot, but if agtoplot is left empty all beads will be plotted, 4 per page.
```{r results, fig.width=13, fig.height=5}
react_summary <- ap_agresults(SBA, samplegroups = factor(SBA$SAMPLES$Diagnosis), groupcolors = 2:6,
  agtoplot = "PPL_HPRR3160218", cofisher = 0.05, shouldpdf = FALSE, useDingbats = FALSE,
  check.names = FALSE)

# Reactivity summary
lapply(react_summary, class)
```

# General data overview and evaluation functions

## Signal overview

ap_overview() plots boxplots per sample and bead, in assay order, sorted by median and sorted by max.
All wells and beads in the input list are included, even if filtering/flagging has been annotated.

```{r overview, fig.width=13, fig.height=13}
ap_overview(SBA, shouldpdf = FALSE, useDingbats = FALSE)
```

## Negative control beads
It is good to see how the negative control beads relate to the rest of the signals.
The function ap_negbeads() plots beeswarms for all samples (without filtering annotation) and 
coloring the points based on the scores. The negative control beads are colored separately, and below the beeswarms 
there are correlation plots between the median sample signal and the negative control signals, colored by the score of the control bead.

In general there is usually a good correlation between the control beads and the medians, 
and the MAD-transformation and scoring adjusts for the different sample background levels.

```{r negbeads, fig.width=13, fig.height=10}
ap_negbeads(SBA, shouldpdf = FALSE, useDingbats = FALSE)
```

# Assay and analysis summary
The function ap_summary() gives you a summary of number of proteins represented, sequence lengths etc.

```{r summary}
ap_summary(SBA, filter = TRUE)
```

## Excel output 
The function ap_excel() writes the data and results to an excel file, where each element corresponds to one sheet.
You select which elements you want to save. 
NB! The function renames and creates some elements, so read the function help documentation for information on how certain sheets are named.

```{r excel, eval=F}
SBA_results <- append(SBA, react_summary)

ap_excel(SBA_results, elements = c("MFI", "MADS", "SCORE", "BINARY_CO",
  "REACTIVITIES", "FISHER", "DIFFERENCES", "ANTIGEN_CUTOFFS", "CUTOFF_KEY",
  "ANTIGENS", "SAMPLES", "COUNT"), filename = "DataOutput.xlsx",
  row.names = TRUE)
```


## Cutoff key
The function ap_cutoffs2image() makes an image of cutoff key, with the corresponding colors.

```{r key, fig.width=10, fig.height=6}
ap_cutoffs2image(cutoffkey = SBA$CUTOFF_KEY, shouldpng = FALSE)
```


## R session
It's always good to save the R session info, which includes R and package version numbers.
```{r session}
sessionInfo()
```

