---
title: "Functions for basic QC and visualization of autoimmunity data (rappp)"
author: "Cecilia Hellström"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Functions for basic QC and visualization of autoimmunity data (rappp)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(rappp)
```

This package includes functions to enable a standardized workflow for quality control and 
some basic analysis and visualizations of autoimmunity data.
It also includes some functions related to peptide design, see the end.

This vignette will demonstrate a basic workflow using the included functions.

# Input data structure

Most functions are based on the below input structure, which is then built upon by the different functions.

```{r}
        SBA <- list(MFI=mfi,
                    COUNT=count,
                    SAMPLES=sampleinfo,
                    BEADS=beads,
                    CT=mfi_ct,
                    FILTERINFO=NULL)
```

The list elements are described in more detail below.

```{r}
head(mfi)
```

```{r}
head(count)
```

```{r}
head(sampleinfo)
```

```{r}
head(beads)
```

```{r}
head(mfi_ct)
```

